---
title: "Getting Started with pmforest"
author: "Kyle Barrett"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=7, 
  fig.height=4
)
```

```{r setup, include=FALSE}
if (interactive()) {
  devtools::load_all()
}
```

```{r load packages, results = 'hide', message=FALSE, warning=FALSE}
library(dplyr)
library(patchwork)
#library(pmforest)
devtools::load_all()
```

<style> body {text-align: justify} </style>

# Introduction/Scope
`pmforest` is a package for producing forest plots: graphical summaries of covariate effects or covariate-dependent predictions in a pharmacometric model. 

Given a dataset of model responses or model parameters with uncertainty at various covariate levels, the package generates plots of confidence intervals reflecting the predicted impact of the covariates. This uncertainty could come from a variety of sources, for example: simulations, bootstrapping, Bayesian posteriors.


<hr>
<br>

# Data Format
To use the forest plot function you will have raw data coming from something like a bunch of simulations. Before plotting, the raw data must first be summarized. The help with this, `pmforest` has a [`summarize_data()`](https://metrumresearchgroup.github.io/pmforest/reference/summarize_data.html) function. The usage of this function is described in the next section, but note that **the user can write their own custom summarization function**, as long as the data it returns conforms to the format described in the "Output Data" section of [`?summarize_data`](https://metrumresearchgroup.github.io/pmforest/reference/summarize_data.html#details).

## Raw data
The raw dataset must be in long format and have the following:

* A numeric column containing the value to be summarized (passed to `value` argument of [`summarize_data()`](https://metrumresearchgroup.github.io/pmforest/reference/summarize_data.html)).
* A character (or factor) column containing the groups to summarize over (passed to `group` argument of [`summarize_data()`](https://metrumresearchgroup.github.io/pmforest/reference/summarize_data.html)).

It can optionally have other grouping columns, described in the `group_level`, `metagroup`, and `replicate` arguments of [`?summarize_data`](https://metrumresearchgroup.github.io/pmforest/reference/summarize_data.html#arguments). Note that specifying a `replicate` column (i.e. a simulation number) will result in a plot with multiple intervals per group level (see [the "Multiple Simulations" vignette](https://metrumresearchgroup.github.io/pmforest/articles/multiple-simulations.html)).

An example dataset that meets these specifications comes with the package (and is shown below). For this data set, 

```{r}
rawData <- readRDS(system.file(file.path("test-data", "plotData.RDS"), package = "pmforest"))
rawData %>%
  select(stat, GROUP, LVL) %>%
  head()
```

In this example, `stat` is the "value", `GROUP` is the "group" column, and `LVL` will be "group_level".

Now we pass this data to [`summarize_data()`](https://metrumresearchgroup.github.io/pmforest/reference/summarize_data.html). This is where the confidence interval and other relevant statistical parameters will be calculated.

```{r}
sumData <- rawData %>%
  summarize_data(
    value = stat,
    group = GROUP,
    group_level = LVL,
    probs = c(0.05, 0.95),
    statistic = "median"
  )
head(sumData)
```

The output data will have three new columns `mid`, `lo`, and `hi`, containing the summarized values (from the data in `value`) corresponding to what was passed to `statistic` (`mid`) and `probs` (`lo`/`hi`). The `group` and `group_level` columns will also be renamed as seen above.

Note that, if a `replicate` column is passed, there will be nine numeric columns (instead of three) output from [`summarize_data()`](https://metrumresearchgroup.github.io/pmforest/reference/summarize_data.html). See [the "Multiple Simulations" vignette](https://metrumresearchgroup.github.io/pmforest/articles/multiple-simulations.html) for an example of this.

### Column names

It's worth noting here that [`plot_forest()`](https://metrumresearchgroup.github.io/pmforest/reference/plot_forest.html) (discussed in the next section) will assume the column names listed above, and will throw an error if they are renamed to anything else. In cases where users would like to reformat these tables for reports, it is recommended to do so _after_ plotting the data.


<hr>
<br>

# Plot Generation

The main plotting function is [`pmforest::plot_forest()`](https://metrumresearchgroup.github.io/pmforest/reference/plot_forest.html). In the examples below we will reference the dataset shown above. Here is a basic example:


```{r}
sumData %>% 
  plot_forest() 
```

<br>

## Formatting Arguments

Below are a few examples showcasing the uses of some of the function arguments, as well as additional formatting techniques using the [patchwork](https://github.com/thomasp85/patchwork) R package. 

<br>

### Basics

You can add a vertical line (at some reference value), and suppress adding the table of confidence intervals.

```{r, fig.width=5}
sumData %>% 
  plot_forest(vline_intercept = 1, 
              annotate_CI = FALSE) 
```


<br>

### Adding group labels

The `summary_labels` argument takes a [`ggplot2::labeller()`](https://ggplot2.tidyverse.org/reference/labeller.html) function. In this example, load a list of covariate labels with the [`yspec` package](https://github.com/metrumresearchgroup/yspec) and pass it through [`ggplot2::as_labeller()`](https://ggplot2.tidyverse.org/reference/as_labeller.html).

```{r}
spec_file <- yspec::ys_load(
  system.file(file.path("test-data", "spec", "analysis3.yml"), package = "pmforest")
)
all_labels <- yspec::ys_get_short(spec_file, title_case = TRUE)
plot_labels <- as_labeller(unlist(all_labels))

sumData %>% 
  plot_forest(vline_intercept = 1, 
              summary_label = plot_labels) 
```
Notice the more informative group labels on the y-axis and the CI table, as compared to the covariate names shown in the previous plots.

### Adding captions and other labels

This example illustrates the use of the other label arguments (`CI_label`, `x_lab`, and `caption`). 

The title at the top of the plot is also pulled from the `metagroup` column of the data, as it is most useful when using `metagroup` to make multiple plots (shown below). Here we just create it with `dplyr::mutate()` (and add a label for it to `plot_labels`).

Some notes: 
* `caption` is a `patchwork` argument. You can change the `caption` formatting using `patchwork` functions (see *Patchwork Annotations* example).
* If `CI_label` is _not_ passed, the value passed to `x_lab` will be used for the table label.
* `text_size` will only apply to `ggplot` layers. 

```{r}
all_labels$CL <- 'CL (L/hr)'
plot_labels <- as_labeller(unlist(all_labels))

sumData %>% 
  mutate(metagroup = "CL") %>%
  plot_forest(vline_intercept = 1, 
              x_lab = "Fraction and 95% CI\nRelative to Reference",
              CI_label = "Median [95% CI]",
              summary_label = plot_labels,
              caption = "This is a patchwork caption.
                         You can add multiple lines.") 
```

<br>

### Multiple covariates (or facets)

Here we pull in a new dataset where the `param` column contains both `VC` and `CL`, where the previous example only contained `CL`, and pass this to the `metagroup` argument of [`summarize_data()`](https://metrumresearchgroup.github.io/pmforest/reference/summarize_data.html).

```{r}
sumData2 <- system.file(file.path("test-data", "plotData2.RDS"), package = "pmforest") %>%
  readRDS() %>%
  summarize_data(
    value = stat,
    group = GROUP,
    group_level = LVL,
    metagroup = param
  )
head(sumData2)
```

The `metagroup` argument doesn't necessarily have to facet across multiple covariates, though this is a common use case. When doing this, you may consider setting `annotate_CI = FALSE` to simplify the plots.

```{r, fig.width=8}
all_labels$CL <- 'CL (L/hr)'
all_labels$V2 <- 'V (L)'
plot_labels <- as_labeller(unlist(all_labels))

sumData2 %>% 
  plot_forest(vline_intercept = 1, 
              x_lab = "Fraction and 95% CI\nRelative to Reference",
              CI_label = "Median [95% CI]",
              summary_label = plot_labels,
              caption = "This is a patchwork caption.
                         You can add multiple lines.",
              annotate_CI = FALSE) 
```

Regardless, you can specify how you want to align the plots using the `nrow` argument or other functions in the `patchwork` R package.

```{r, fig.height=8}
sumData2 %>% 
  plot_forest(vline_intercept = 1, 
              x_lab = "Fraction and 95% CI\nRelative to Reference",
              CI_label = "Median [95% CI]",
              summary_label = plot_labels,
              caption = "This is a patchwork caption.
                         You can add multiple lines.",
              nrow = 2) 
```

<br>


### Shaded Interval
You can add a shaded interval over a specified range. It is important to note that the x-axis is automatically determined based on your input data. The specified interval may therefore be cut off, as the x limits do not adjust based on the specified interval. You can correct for this by specifying the `x_breaks` and `x_limit` arguments.

```{r}
sumData %>% 
  mutate(metagroup = "CL") %>%
  plot_forest(vline_intercept = 1, 
              x_lab = "Fraction and 95% CI\nRelative to Reference",
              CI_label = "Median [95% CI]",
              summary_label = plot_labels,
              caption = "This is a patchwork caption.
                         You can add multiple lines.",
              shaded_interval = c(0.8,1.25),
              x_breaks = c(0.4,0.6, 0.8, 1, 1.2, 1.4, 1.6), 
              x_limit = c(0.6,1.45)) 
```

<br>

### Patchwork Annotations
The individual layers of the returned object are defined using `ggplot`, but the final plot is assembled using `patchwork`.

Thus you can add additional `patchwork` themes/annotations to the returned object, but cannot alter the ggplot layers. Themes can be added using the `&` operator, and follow the same convention as `ggplot` themes. In this example we change the formatting of the `caption` argument.

```{r, fig.height=7, fig.width=8}
clp <- sumData %>% 
  mutate(metagroup = "CL") %>%
  plot_forest(vline_intercept = 1, 
              x_lab = "Fraction and 95% CI\nRelative to Reference",
              CI_label = "Median [95% CI]",
              summary_label = plot_labels,
              caption = "This is a patchwork caption.
                         You can add multiple lines.") 

wrap_plots(clp, grid::textGrob('Text on right side'), widths = c(4,1)) +
  plot_annotation(
    title = 'Here is a title',
    subtitle = 'Here is a subtitle',
    caption = 'Overrides plot_forest caption argument',
    tag_levels = c('A', '1'), tag_prefix = 'Fig. ', tag_sep = '.', tag_suffix = ':',
    theme = theme(plot.title = element_text(size = 18)))  & 
  theme(plot.caption = element_text(size = 12),
        plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 8, hjust = 0, vjust = 0))
```

<br>


### Full Forest Plot Example:

```{r, fig.height=7, fig.width=8}
clp <- sumData %>% 
          plot_forest(shaded_interval = c(0.8,1.25),
                      summary_label = plot_labels,
                      text_size = 3.5,
                      vline_intercept = 1,
                      x_lab = "Fraction and 95% CI \nRelative to Reference",
                      CI_label = "Median [95% CI]",
                      caption = "The shaded area corresponds 
                                 to the interval (0.8, 1.25)",
                      plot_width = 9, # out of 12
                      x_breaks = c(0.4,0.6, 0.8, 1, 1.2, 1.4,1.6), 
                      x_limit = c(0.6,1.45),
                      annotate_CI = TRUE)
                   
clp +
    plot_annotation(
          title = 'Here is a title',
          subtitle = 'Here is a subtitle',
          theme = theme(plot.title = element_text(size = 18))) & 
    theme(plot.caption = element_text(size = 11))
```

