---
title: "Multiple Simulations"
author: "Kyle Barrett"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{multiple-simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=8, 
  fig.height=5.5
)
```

```{r setup, include=FALSE}
if (interactive() || isTRUE(getOption('knitr.in.progress'))) {
  devtools::load_all()
}
```

```{r load packages, results = 'hide', message=FALSE, warning=FALSE}
library(dplyr)
library(patchwork)
library(pmforest)
```

<style> body {text-align: justify} </style>

### Introduction
In some instances, it may be useful to present the variability of each summary statistic (i.e. median, 5th and 95th percentile). In order to accomplish this you must have a column denoting the replicate/simulation number. This column is then passed to the `replicate` argument. An example dataset comes with the package:

```{r}
rawData <- readRDS(system.file(file.path("test-data", "plotDataEXP.RDS"), package = "pmforest"))
rawData %>%
  select(nsim, stat, GROUP, LVL) %>%
  head()
```

In this example, `nsim` is the "replicate" identifier, while `stat` is the "value", `GROUP` is the "group" column, and `LVL` will be "group_level". Notice in our example we have 50 replicates and 1,800 rows for each replicate.

```{r}
count(rawData, nsim)
```

<br>

# Summarize data

When summarizing the data, users will have to specify the `replicate` column. This will output the same data structure described in the "Summarize data" section of [the "Getting Started" vignette](https://metrumresearchgroup.github.io/pmforest/articles/getting-started.html#summarize-data), however it will now have _nine_ numeric columns instead of three.

```{r}
sumData <- 
  summarize_data(
    rawData, 
    value = stat,
    group = GROUP,
    group_level = LVL,
    replicate = nsim
  )
head(sumData)
```

These nine columns (`mid_mid`, `mid_lo`, `mid_hi`, plus three more each for `lo_*` and `hi_*`) contain the summarized values. In this case, the `mid_mid`, `lo_mid`, and `hi_mid` columns correspond to the values of the major interval (i.e. the big lines and central data point on the plots we are about to create) and the `*_lo` and `*_hi` columns correspond to the values for each of the three minor intervals (i.e. the small lines on the plots).

As before, `plot_forest` will assume the grouping column names shown above, and will throw an error if they are renamed to anything else.

# Create plots

You can now pass this data directly to [`pmforest::plot_forest()`](https://metrumresearchgroup.github.io/pmforest/reference/plot_forest.html). The function will recognize the additional columns and automatically draw the minor intervals.

Note: these plots can be customized and labelled as shown in [the "Getting Started" vignette](https://metrumresearchgroup.github.io/pmforest/articles/getting-started.html#formatting-arguments), but for clarity and simplicity, we just show the basic forest plots below.

```{r}
plot_forest(sumData, vline_intercept = 1)
```

### Jitter minor interval

In cases where the additional confidence intervals overlap, you can vertically jitter them via the `jitter_reps` argument (logical). An example can be seen below (note that the data is different than above):

```{r}
system.file(file.path("test-data", "plotDataEXP.RDS"), package = "pmforest") %>%
  readRDS() %>%
  mutate(stat = stat + (nsim/600)) %>% # making variability larger to demonstrate jitter
  summarize_data(
    value = stat,
    group = GROUP,
    group_level = LVL,
    replicate = nsim
  ) %>%
  plot_forest(
    jitter_reps = TRUE,
    vline_intercept = 1
  )
```

